---
title: "AERONET_PM_continuous_CA"
author: "Meytar Sorek-Hamer"
date: "January 18, 2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Libraries
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(purrr)
library(tidyr)
library(ggplot2)
library(data.table)
library(dplyr)
library(Hmisc)
library(rlist)
```
## Load the data
# List Directories
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
#AERONET data
aer_aot_dir <- "C:/Users/msorekha/Documents/Projects/JPL/MISR/Data/AERONET/Stage2/Cont_AOT"
aer_sda_dir <- "C:/Users/msorekha/Documents/Projects/JPL/MISR/Data/AERONET/Stage2/Cont_SDA"
aer_inv_dir <- "C:/Users/msorekha/Documents/Projects/JPL/MISR/Data/AERONET/Stage2/Cont_INV"
#aer_inv_pfn_dir <- "C:/Users/msorekha/Documents/Projects/JPL/MISR/Data/AERONET/INV/INV_PFN"
res_dir <- "C:/Users/msorekha/Documents/Projects/JPL/MISR/plots/Stage2/"

#EPA
epa_daily_dir <- "C:/Users/msorekha/Documents/Projects/JPL/MISR/Data/EPA/Data/daily/spec"
```

# List files per directory
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
#Aeronet AOD files
aer_aod <- list.files(path=aer_aot_dir,full.names = TRUE)
#Aeronet SDA files
aer_sda <- list.files(path=aer_sda_dir,full.names = TRUE)
#Aeronet inversion  files
aer_inv <- list.files(path=aer_inv_dir,full.names = TRUE)
#aer_inv_pfn <- list.files(path=aer_inv_pfn_dir,full.names = TRUE)

#EPA daily dust concentrations
epa_daily <- list.files(path=epa_daily_dir,pattern="*.csv",full.names = TRUE)
```

# Read EPA dust data
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
l<-list()
ind=1
for (I in epa_daily){
  l[[ind]] <- fread(I)
  ind=ind+1
}
dust <- rbindlist(l) #all stations in the USA, 2000-2018, 4,688,508
setnames(dust,"CBSA.Name","Site_Name")

setnames(dust,"date","Date")
dust$Date <- as.Date(dust$Date )

dust$Site_Name <- 
      dust$Site_Name %>% 
      strsplit("\\,") %>% 
      sapply("[", 1) 

dust[dust$Site_Name=="Visalia-Porterville",]$Site_Name <- "Visalia"

dust <- filter(dust,!is.na(Date))#5,497

setDT(dust)
#exclude days with negative levels
dust[dust < 0] <- NA#5,495 obs

#calculate DUST
#Dust = 2.2Al +  2.49Si +  1.63Ca +  1.94Ti +  2.42Fe
dust$Dust <- 2.2 * dust$Al +
             2.49 * dust$Si +
             1.63 * dust$Ca +
             1.94 * dust$Ti + 
             2.42 * dust$Fe
  
dust <- dust[,-c("latitude" , "longitude", "POC","temp_mean","temp_min" , "temp_max",  "atmpres"  )]

dust%>%group_by(Site_Name)%>%summarise(N=n(),PM25=mean(PM25,na.rm=T),dust=mean(Dust,na.rm=T),N_date=length(unique(Date)),MinDate=min(Date,na.rm=T),MaxDate=max(Date,na.rm=T))

#save
#write.csv(dust,"C:/Users/msorekha/Documents/Projects/JPL/MISR/Data/WDB/EPA.dust.csv")
```

#Read AERONET AOD-spec data
```{r message=FALSE, warning=FALSE, paged.print=FALSE,include=FLASE}
#AOD
l<-list()
for (I in aer_aod){
  stn.name <- 
      I %>% 
      strsplit("\\.") %>% 
      lapply(rev) %>% 
      sapply("[", 2) %>% 
      strsplit("\\d_")%>%
      lapply(rev) %>% 
      sapply("[", 1)
  l[[stn.name]] <- fread(I,skip=6,fill=TRUE)
}
aer.aod <- rbindlist(l,fill=TRUE,idcol=TRUE) 
setnames(aer.aod,".id","Site_Name")
setnames(aer.aod,"Date(dd:mm:yyyy)","Date")
setnames(aer.aod,"Time(hh:mm:ss)","hour")
aer.aod$Date <- as.Date(aer.aod$Date,format="%d:%m:%Y" )

#change site names
aer.aod$Site_Name[aer.aod$Site_Name=="DRAGON_Bakersfield"] <- "Bakersfield"
aer.aod$Site_Name[aer.aod$Site_Name=="Fresno_2"] <- "Fresno"
aer.aod$Site_Name[aer.aod$Site_Name=="DRAGON_Visalia"] <- "Visalia"

#remove duplicate columns
coll <- which(aer.aod[,!duplicated(colnames(aer.aod))])
aer.aod <- aer.aod[,..coll]

#Daily averages from hourly data
aer.aod.day <- aer.aod %>%
          group_by(Site_Name,Date) %>%
          summarise_all(.,mean) #4938; 3978 days

setDT(aer.aod.day)
aer.aod.day[aer.aod.day<0]<-NA

#interpolate aeronet AOD to 550nm
aer.aod.day$AOD_550 <- as.numeric(aer.aod.day$AOD_440nm)*(550/440)^(-(aer.aod.day$`440-675_Angstrom_Exponent`))

vars <- fread(("C:/Users/msorekha/Documents/Projects/JPL/MISR/Data/WDB/vars.type.spec.csv"))

setDF(aer.aod.day)
aod <-aer.aod.day[,names(aer.aod.day) %in% vars$Var1 ]
aod <- aod %>% select(-contains("N["))%>% select(-contains("Flux"))%>% select(-contains("(Dobson)"))%>% select(-contains("Exact"))
aod <- cbind(aer.aod.day[,1:2],aod)

aod%>%group_by(Site_Name)%>%summarise(N=n(),AOD=mean(AOD_550,na.rm=T),N_date=length(unique(Date)),MinDate=min(Date,na.rm=T),MaxDate=max(Date,na.rm=T))
```
#Merge ARONET and dust hourly based daily mean
```{r}

#Merge ARONET and dust daily mean
aer.coll.d <- merge(aod,dust,by=c("Site_Name","Date")) #1,375 obs X 76 vars

setDF(aer.coll.d )
#exclude vars with less than 80% availability
wdb.f <- aer.coll.d[,colSums(is.na(aer.coll.d))/dim(aer.coll.d)[1]<0.8]#1,375 obs X 58 vars

#Save files
#write.csv(aer.aod,paste0(res_dir,"aer.aod.cont.csv"))
#write.csv(wdb.x,paste0(res_dir,"AODonly.dust.CSN.csv"))

wdb.x%>%
  group_by(Site_Name)%>%
  summarise(N=n(),dust=mean(Dust,na.rm=T),PM25=mean(PM25,na.rm=T),AOD=mean(AOD_550,na.rm=T),N_date=length(unique(Date)))
```

#raw correlation
```{r}
 require(Metrics)
wdb.x%>%
            group_by(Site_Name)%>%
            summarise(N=n(),r=cor(PM25,AOD_550,use='na.or.complete'), r2=cor(PM25,AOD_550,use='na.or.complete')^2,rmse=rmse(PM25, (lm(PM25~AOD_550,data=.))$fitted.values))

```


#Read AERONET INV data 
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
l<-list()
for (I in aer_inv){
  stn.name <- 
      I %>% 
      strsplit("\\_") %>% 
      lapply(rev) %>% 
      sapply("[", 1) %>% 
      strsplit("\\.")%>%
      lapply(rev) %>% 
      sapply("[", 2)
  l[[stn.name]] <- fread(I,skip=6,fill=TRUE)
}
aer.inv<- setDT(rbindlist(l,idcol=TRUE)) #daily 2679x242
setnames(aer.inv,".id","Site_Name")
setnames(aer.inv,"Date(dd:mm:yyyy)","Date")
setnames(aer.inv,"Time(hh:mm:ss)","hour")
aer.inv$Date <- as.Date(aer.inv$Date, format="%d:%m:%Y")    

#change site names
aer.inv$Site_Name[aer.inv$Site_Name=="2"] <- "Fresno"

#merge data to collocated aeronet AOD dataset
aot.dust.inv <- merge(aer.coll.d,aer.inv,by=c("Site_Name","Date"),all=TRUE)#4,450 x 276 vars

setDF(aot.dust.inv )
#exclude vars with less than 60% availability
wdb.f <- aot.dust.inv[,colSums(is.na(aot.dust.inv))/dim(aot.dust.inv)[1]<0.4]#4,450 obs X 276 vars

vars <- fread(("C:/Users/msorekha/Documents/Projects/JPL/MISR/Data/WDB/vars.type.spec.csv"))
new <- data.frame(Var1="AOD_550",group=NA)
vars<-rbind(vars,new)
setDF(wdb.f)
wdb.x <- wdb.f[,names(wdb.f) %in% vars$Var1 ]#4,450 obs X 63 vars

wdb.x <- cbind(wdb.f[,1:2],wdb.x)

#Save files
# write.csv(wdb.x,paste0(res_dir,"AOD.dust.inv.CONT.csv"))

```

#Read AERONET SDA data
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
#SDA
l<-list()
for (I in aer_sda){
  stn.name <- 
      I %>% 
      strsplit("\\.") %>% 
      lapply(rev) %>% 
      sapply("[", 2) %>% 
      strsplit("\\d_")%>%
      lapply(rev) %>% 
      sapply("[", 1)
  if (I==aer_sda[1]){
  x1 <- fread(I,skip=6,fill=TRUE)
  l[[stn.name]] <- x1[,-c("V35","V36")]
  } else {
   l[[stn.name]] <- fread(I,skip=6,fill=TRUE) 
  }
}
aer.sda <- setDT(rbindlist(l,idcol=TRUE)) #2671x35
setnames(aer.sda,".id","Site_Name")
setnames(aer.sda,"Date_(dd:mm:yyyy)","Date")
setnames(aer.sda,"Time_(hh:mm:ss)","hour")
aer.sda$Date <- as.Date(aer.sda$Date, format="%d:%m:%Y")    

#change site names
aer.sda$Site_Name[aer.sda$Site_Name=="DRAGON_Bakersfield"] <- "Bakersfield"
aer.sda$Site_Name[aer.sda$Site_Name=="Fresno_2"] <- "Fresno"

 #merge data to collocated aeronet AOD dataset
aot.dust.inv.sda <- merge(aot.dust.inv,aer.sda,by=c("Site_Name","Date"))#3859
aot.dust.sda <- merge(aer.coll,aer.sda,by=c("Site_Name","Date"))#5831

#Save files
#write.csv(aot.dust.sda,paste0(aer_aot_dir,"aot.dust.sda.csv"))
#write.csv(aot.dust.inv.sda ,paste0(aer_aot_dir,"aot.dust.inv.sda.csv"))
```
#SDA has very poor availability do in the first stage wont be included in analysis

#Histogram of all variables
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
##AERONET AOD - all DRAGON stations 
#Convert all character variables to numeric variables
aer.aod <-aer.coll.inv

#1:68 = AOD; 70:104 = sda; 106:343 = inv
aer.aod <- aer.aod[,c(1:20,39:44,72:83,126:166,190:201)]
#3:18 = AOD; 19:36 = sda; 37:89 = inv
#remove 'hour' and 'julian day')
aer.aod <- aer.aod[,-c(3:4)]

#create source matrix
source.aod<-setDF(aer.aod[,-c(1:2)])
source.aod[source.aod=="N/A"]<-NA
source.aod[!is.na(source.aod)]<-"1"
source.aod <- cbind(aer.aod[,c(1:2)],source.aod)

source.aod.0<-source.aod[,1:2]
source.aod.1<-source.aod[,3:18]; source.aod.1[source.aod.1=='1']<-"AOD"
source.aod.2<-source.aod[,19:36]; source.aod.2[source.aod.2=='1']<-"SDA"
source.aod.3<-source.aod[,37:89]; source.aod.3[source.aod.3=='1']<-"INV"

source<-cbind(source.aod.0,source.aod.1,source.aod.2,source.aod.3)

#plot Data availability
source.melt <- melt(source,id.vars=c("Site_Name","Date"))
source.melt <- na.omit(source.melt)

#plot availability bar chart
ggplot(data=source.melt,aes(variable)) +
    geom_bar(aes(fill=value,weight=1/120))+
  ylab("% Availability")+xlab("")+
    facet_wrap(~ Site_Name)+ 
    theme(axis.text.x = element_blank())

ggsave("C:/Users/msorekha/Documents/Projects/JPL/MISR/plots/Stage2/availability_bystation.png",width=15, height=7)

ggplot(data=source.melt,aes(variable)) +
    geom_bar(aes(fill=type,weight=1/1014))+
    ylab("% Availability")+xlab("")+
    theme(axis.text.x = element_text(size=rel(0.65),angle=45,hjust=1))

ggsave("C:/Users/msorekha/Documents/Projects/JPL/MISR/plots/Stage2/availability_all.png",width=15, height=7)

#list vars by type
aod.vars<-names(source.aod.1)
sda.vars<-names(source.aod.2)
inv.vars<-names(source.aod.3)
vars<-qpcR:::cbind.na(aod.vars, sda.vars,inv.vars)
write.csv(vars,"C:/Users/msorekha/Documents/Projects/JPL/MISR/plots/Stage2/vars_type.csv")

#count Data availability
count<-source.melt%>%group_by(Site_Name)%>%summarise(sum.aod=sum(type=='AOD'),sum.sda=sum(type=='SDA'),sum.inv=sum(type=='INV'))
write.csv(count,"C:/Users/msorekha/Documents/Projects/JPL/MISR/plots/Stage2/vars_avail.csv")

#plot histogram per variable
#convert all vars to numeric format
time <- aer.aod[,c(1:2)]
aer.aod.ch<-aer.aod[,-c(1:2)] %>%
    keep(is.character)%>%
aer.aod.ch[] <- lapply(aer.aod.ch, function(x) as.numeric(as.character(x)))
aer.aod.num <- aer.aod[,-c(1:2)] %>%
    keep(is.numeric)
aer.aod <- cbind(time, aer.aod.ch,aer.aod.num)
sum.vars <- colSums(is.na(aer.aod))/dim(aer.aod)[1]
sum1<-c(names(which(sum.vars==1)))

  aer.aod[,-..sum1]%>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram(binwidth = 0.1, fill="black", col="grey")+
    theme(axis.text = element_text(size=5),
          strip.text.x = element_text(size = 7))

  ggsave("C:/Users/msorekha/Documents/Projects/JPL/MISR/plots/Stage2/hist_all.png",width=15, height=7)

  #Convert all character variables to numeric variables
aer.sda <- aer.sda[,c(1:15)]

#plot histogram per variable
aer.sda%>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram(binwidth = 0.1, fill="black", col="grey")

##EPA dust Conc
#plot histogram per variable
dust.coll%>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram(bins=100,fill="black", col="grey")
```
 
## Availability Plot
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
#Check availability per station
stn.name <- unique(aeronet.m$Site_Name)
date <- unique(aeronet.m$Date)#120 out of 150

avail.by.stn <- aeronet.m %>% 
  group_by(Site_Name)%>%
  summarise(N=n())
#save file - availability by station
write.csv(avail.by.stn,paste0(aer_aot_dir,"avail.by.stn.csv"))
  
  ###
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(date,value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram(binwidth = 0.1, fill="black", col="grey")

df <- data.frame(year,group,value)
df$value[df$value==1] <- NA

no_na_df <- df[complete.cases(df), ]

ggplot(no_na_df, aes(x=year, y = group)) + 
    geom_errorbarh(aes(xmax = year, xmin = year), size = 2)
```


#save wdb 
```{r}
write.csv(aot.dust.inv,"C:/Users/msorekha/Documents/Projects/JPL/MISR/Data/WDB/wdb_pm.cont.csv")

```

#extract list of variables from 352
```{r}
vars.list <- fread("C:/Users/msorekha/Documents/Projects/JPL/MISR/Data/WDB/vars.type.spec.csv")
vars.list <- as.list(vars.list)  
setDT(aot.dust.inv)
wdb.ex <- aot.dust.inv[,which(colnames(aot.dust.inv) %in% vars.list[[1]]),with=FALSE]#14,465X81
#wdb.ex[wdb.ex=='N/A']<-NA
wdb.ex[wdb.ex=='-999']<-NA
wdb.ex <- cbind(aot.dust.inv[,1:2],wdb.ex)

#remove duplicate rows
row <- unique(wdb.ex)


#exclude vars with less than 90% availability
wdb.f <- wdb.ex[,(colSums(is.na(wdb.ex)/dim(wdb.ex)[1]))<0.1,with=FALSE]
write.csv(wdb.f,"C:/Users/msorekha/Documents/Projects/JPL/MISR/Data/WDB/wdb.cont.csv")
write.csv(names(wdb.f),"C:/Users/msorekha/Documents/Projects/JPL/MISR/Data/WDB/wdb.cont.vars.csv")
```

#